---
title: "MVA-Final Project"
author: "Leaf Dataset Group"
date: "11/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
# install.packages('comprehenr')
library(comprehenr)
library(KernSmooth)
library(ellipse)
library(MVA)
library(CCA)  
# comment
 
full_df = read.csv('https://raw.githubusercontent.com/iniadewumi/MVA-Dataset-Exploration/master/MVA/datasets/leaf/leaf.csv')

k = c("1" = "Quercus suber", "2" = "Salix atrocinerea", "3" = "Populus nigra", "4" = "Alnus sp", 
      "5" = "Quercus robur", "6" = "Crataegus monogyna", "7" = "Ilex aquifolium", "8" = "Nerium oleander",
      "9" = "Betula pubescens", "10" = "Tilia tomentosa", "11" = "Acer palmaturu", "12" = "Celtis sp", 
      "13" = "Corylus avellana", "14" = "Castanea sativa", "15" = "Populus alba", "16" = "Acer negundo", 
      "17" = "Taxus bacatta", "18" = "Papaver sp", "19" = "Polypodium vulgare", "20" = "Pinus sp", 
      "21" = "Fraxinus sp", "22" = "Primula vulgaris", "23" = "Erodium sp", "24" = "Bougainvillea sp", 
      "25" = "Arisarum vulgare", "26" = "Euonymus japonicus", "27" = "Ilex perado ssp azorica", 
      "28" = "Magnolia soulangeana", "29" = "Buxus sempervirens", "30" = "Urtica dioica", 
      "31" = "Podocarpus sp", "32" = "Acca sellowiana", "33" = "Hydrangea sp", "34" = "Pseudosasa japonica", 
      "35" = "Magnolia grandiflora", "36" = "Geranium sp", "37" = "Aesculus californica", 
      "38" = "Chelidonium majus", "39" = "Schinus terebinthifolius", "40" = "Fragaria vesca")

cols = c('Class', 'Specimen_Number', 'Eccentricity', 'Aspect_Ratio', 'Elongation', 'Solidity', 
         'Stochastic_Convexity', 'Isoperimetric_Factor', 'Maximal_Indentation_Depth', 'Lobedness', 
         'Average_Intensity', 'Average_Contrast', 'Smoothness', 'Third_Moment', 'Uniformity', 
         'Entropy')


colnames(full_df) = cols


for (i in 1:nrow(full_df)) {
  name = as.character(k[full_df[i, "Class"]])
  full_df[i, "Sci_Name"] = name
}


full_df = as.data.frame(full_df)
leaf_df = full_df[-c(1,2,17) ]
leaf_df
```


##MISSING DATA CHECK
*We will check for NA values, and see if there are any zero values, which might suggest that the value might have been missing value. *
*The we will check for negative values as well, which might affect analysis in the future*
*Next we conduct a basic summary on the data-frame.*

```{r}
cat("Count of NAs:", sum(is.na(leaf_df)), "\n")
cat("Count of 0 values", sum(leaf_df[leaf_df==0]), "\n")
cat("Cosunt of Negative values:", sum(leaf_df[leaf_df<0]), "\n")

cat("\n\n")
summary(leaf_df)
```

##OUTLIERS CHECK
```{r}
bvbox(leaf_df)

```

##DATASET CORRELATION AND COVARIANCE 
```{r}

leaf_df.cor = cor(leaf_df)
leaf_df.cov = cov(leaf_df)

cat("Covariance Matrix\n\n")
leaf_df.cov

cat("\n\nCorrelation Matrix\n\n")
leaf_df.cor

```

##FINDINGS FROM CORRELATION AND COVARIANCE ANALYSIS

```{r}

```

##DENSITY ANALYSIS

```{r}


#Use list comprehension to calculate the bw for each columns in the dataset
bw = to_vec(for(col in colnames(leaf_df)) dpik(as.matrix(leaf_df[col])))


#Calculate the density for each column
leaf_df.mdens <- bkde2D(leaf_df, bandwidth = bw)



# plot(leaf_df)
# contour(x = leaf_df.mdens$x1, y = leaf_df.mdens$x2, z = leaf_df.mdens$fhat, add = TRUE)


for( col in colnames(leaf_df)){
   
  bw = dpik(as.matrix(leaf_df[col]))
  
  den = density(as.matrix(leaf_df[col]), bw=bw, kernel = "gaussian")
  plot(den, main=col)
}


```

##NOTE: Multidimension Density Plot is unreadable. 



##PCA Analysis
```{r}

leaf_df.pca_cor = princomp(leaf_df, cor = TRUE)
summary(leaf_df.pca_cor, loadings=T)

```


##MDS Analysis
```{r}
leaf.dist = dist(leaf_df)

leaf_df.mds = cmdscale(leaf.dist, eig = TRUE)
leaf_df.mds

```

##CCA Analysis
```{r}
# setup indices for two groups
G1 <- 3:9
G2 <- 10:16

leaf_df.cca <- cc(leaf_df[,G1], leaf_df[,G2])

leaf_df.cca

```

##Hierarchical Clustering
```{r}


hcs <- hclust(leaf.dist, "single")
plot(hcs, main = "Single Linkage HC Dendrogram")

```


```{r}
print("hello")
```

