---
title: "MVA-Final Project"
author: "Leaf Dataset Group"
date: "11/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
# install.packages('comprehenr')
# install.packages('ggfortify')
library(comprehenr)
library(KernSmooth)
library(ellipse)
library(MVA)
library(CCA)
library(dplyr)
# comment
library(ggfortify)





plot.wgss <- function(mydata, maxc) {
  wss <- numeric(maxc)
  for (i in 1:maxc) {
    wss[i] <- kmeans(mydata,iter.max = 100, centers = i, nstart = 10)$tot.withinss
  }
  plot(1:maxc, wss,
    type = "b",
    xlab = "Number of Clusters",
    ylab = "Within Groups Sum of Squares",
    main = "Scree Plot"
  )
}




full_df = read.csv(
  'https://raw.githubusercontent.com/iniadewumi/MVA-Dataset-Exploration/master/MVA/datasets/leaf/leaf.csv'
)


k = c(
  "1" = "Quercus suber",
  "2" = "Salix atrocinerea",
  "3" = "Populus nigra",
  "4" = "Alnus sp",
  "5" = "Quercus robur",
  "6" = "Crataegus monogyna",
  "7" = "Ilex aquifolium",
  "8" = "Nerium oleander",
  "9" = "Betula pubescens",
  "10" = "Tilia tomentosa",
  "11" = "Acer palmaturu",
  "12" = "Celtis sp",
  "13" = "Corylus avellana",
  "14" = "Castanea sativa",
  "15" = "Populus alba",
  "16" = "Acer negundo",
  "17" = "Taxus bacatta",
  "18" = "Papaver sp",
  "19" = "Polypodium vulgare",
  "20" = "Pinus sp",
  "21" = "Fraxinus sp",
  "22" = "Primula vulgaris",
  "23" = "Erodium sp",
  "24" = "Bougainvillea sp",
  "25" = "Arisarum vulgare",
  "26" = "Euonymus japonicus",
  "27" = "Ilex perado ssp azorica",
  "28" = "Magnolia soulangeana",
  "29" = "Buxus sempervirens",
  "30" = "Urtica dioica",
  "31" = "Podocarpus sp",
  "32" = "Acca sellowiana",
  "33" = "Hydrangea sp",
  "34" = "Pseudosasa japonica",
  "35" = "Magnolia grandiflora",
  "36" = "Geranium sp",
  "37" = "Aesculus californica",
  "38" = "Chelidonium majus",
  "39" = "Schinus terebinthifolius",
  "40" = "Fragaria vesca"
)
cols = c(
  'Class',
  'Specimen_Number',
  'Eccentricity',
  'Aspect_Ratio',
  'Elongation',
  'Solidity',
  'Stochastic_Convexity',
  'Isoperimetric_Factor',
  'Maximal_Indentation_Depth',
  'Lobedness',
  'Average_Intensity',
  'Average_Contrast',
  'Smoothness',
  'Third_Moment',
  'Uniformity',
  'Entropy'
)


colnames(full_df) = cols



for (i in 1:nrow(full_df)) {
  name = as.character(k[full_df[i, "Class"]])
  full_df[i, "Sci_Name"] = name
}



outliers <- function(x) {
  Q1 <- quantile(x, probs = .10)
  Q3 <- quantile(x, probs = .90)
  iqr = Q3 - Q1
  upper_limit = Q3 + (iqr * 1.5)
  lower_limit = Q1 - (iqr * 1.5)
  
  if (upper_limit == lower_limit) {
    FALSE
  }
  x > upper_limit | x < lower_limit
}


remove_outliers <- function(df, cols = names(df)) {
  for (col in cols) {
    df <- df[!outliers(df[[col]]), ]
  }
  df
}



test = full_df %>% group_split(Sci_Name)
new_df = df <- data.frame()


for (i in 1:length(test)) {
  df = as.data.frame(test[[i]])[-c(1, 2, 17)]
  out_df = remove_outliers(df, as.vector(colnames(df)))
  new_df = rbind(new_df, out_df)
}



c = 0


for (i in 1:length(test)) {
  df = as.data.frame(test[[i]])
  # cat("\n", df$Class[1], "\n")
  df = df[-c(1, 2, 17)]
  
  for (name in colnames(df)) {
    col = df[, name]
    m = mean(col)
    sdv = sd(col)
    
    for (val in col) {
      if ((val > m + sdv * 3) || (val < m - sdv * 3)) {
        # cat("\n", name, ":", val, "[", m, sdv * 3, "]")
        # cat("\n", val - m + sdv * 3, "\n\n")
        c = c + 1
      }
    }
  }
  
}

```



```{r}
check = full_df[full_df$Class==14,]
check[check$Lobedness == 0.17738 ,]
```


##Exploring leaf 28
```{r}
leaf_28 = full_df[full_df$Class=="28",][-c(1, 2, 17)]
leaf_28
for (col in colnames(leaf_28)) {
  x = leaf_28[,"Stochastic_Convexity"]
  Q1 <- quantile(x, probs=.25)
  Q3 <- quantile(x, probs=.75)
  iqr = Q3-Q1
 upper_limit = Q3 + (iqr*3)
 lower_limit = Q1 - (iqr*3)
 
  if (upper_limit == lower_limit){
    next
  }
  c = 0
  for(x in x > upper_limit | x < lower_limit){
    if(x==TRUE){
        
       cat("\n\n", col, "\n")
       cat(upper_limit, lower_limit)
       c = c +1
    }    
  } 
}
```

##DATASETS
```{r}
# agg_df = aggregate(full_df, list(full_df$Class), mean)
# 
# rownames(agg_df) = agg_df$Sci_Name
# 
# agg_df = agg_df[-c(1,2,3,18) ]
leaf_df = full_df[-c(156),]
num_leaf_df = leaf_df[, -c(1,2,17)]





num_leaf_df.shape_feat = num_leaf_df[c(1:7)]
num_leaf_df.text_feat = num_leaf_df[c(8:14)]

agg = aggregate(leaf_df, list(leaf_df$Class), mean)
for (i in 1:nrow(agg)) {
  name = as.character(k[agg[i, "Class"]])
  agg[i, "Sci_Name"] = name
}


rownames(agg) = agg$Sci_Name


num_agg = agg[, -c(1,2, 3,18)]


agg.shape_feat = num_agg[c(1:7)]
agg.text_feat = num_agg[c(8:14)]


```





##MISSING DATA CHECK
*We will check for NA values, and see if there are any zero values, which might suggest that the value might have been missing value. *
*The we will check for negative values as well, which might affect analysis in the future*
*Next we conduct a basic summary on the data-frame.*

```{r}
cat("Count of NAs:", sum(is.na(num_leaf_df)), "\n")
cat("Count of 0 values", sum(num_leaf_df[num_leaf_df==0]), "\n")
cat("Count of Negative values:", sum(num_leaf_df[num_leaf_df<0]), "\n")

cat("\n\n")
summary(num_leaf_df)
```

##OUTLIERS CHECK
Since we had multiple columns, it was difficult plotting the BV plot with the right variables, so, we decided to do a column by column check on our dataset. We got the  Mean and Standard Deviation of each column, then we checked each row for 3 standard deviations from the mean, and they all fit within 3 standard deviations



But when we used interquartile range, outliers were found. 



##DATASET CORRELATION AND COVARIANCE 
```{r}

num_leaf_df.cor = cor(num_leaf_df)
num_leaf_df.cov = cov(num_leaf_df)

cat("Covariance Matrix\n\n")
num_leaf_df.cov

cat("\n\nCorrelation Matrix\n\n")
num_leaf_df.cor

library(corrplot)
corrplot(num_leaf_df.cor, method = "number", number.cex = 0.5,type = "lower", order = "hclust", col = brewer.pal(n = 7, name = "RdBu"))
```

##FINDINGS FROM CORRELATION AND COVARIANCE ANALYSIS



##DENSITY ANALYSIS

```{r}


#Use list comprehension to calculate the bw for each columns in the dataset
bw = to_vec(for(col in colnames(num_leaf_df)) dpik(as.matrix(num_leaf_df[col])))


#Calculate the density for each column
num_leaf_df.mdens <- bkde2D(num_leaf_df, bandwidth = bw)



# plot(num_leaf_df)
# contour(x = num_leaf_df.mdens$x1, y = leaf_df.mdens$x2, z = leaf_df.mdens$fhat, add = TRUE)


for( col in colnames(num_leaf_df)){
  
  bw = dpik(as.matrix(num_leaf_df[col]))
  
  den = density(as.matrix(num_leaf_df[col]), bw=bw, kernel = "gaussian")
  plot(den, main=col)
}


```

##NOTE: Multidimension Density Plot is unreadable. 

##Histogram for Lobedness and Elongation
```{r}
techst <- hist(num_leaf_df$Lobedness)
num_leaf_df$Lobedness
plot(techst, main="Lobedness Histogram")
Elong <- hist(num_leaf_df$Elongation)
num_leaf_df$Elongation
plot(Elong, main="Elongation Histogram")
```

##Scatterplot comparing Lobedness and Elongation
```{r}
plot(num_leaf_df$Elongation,num_leaf_df$Lobedness ,
     xlab = "Average Elongation",
     ylab = "Average Lobedness ",
     pch = 20, main="Lobedness and Elongation Scatterplot")
```


##PCA Analysis
```{r}

library(devtools)
install_github("vqv/ggbiplot")

library(ggbiplot)

ggbiplot(num_leaf_df.pca_cor)
ggbiplot(num_leaf_df.pca_cor,ellipse=TRUE,circle=TRUE, abels=rownames(leaf_df))


num_leaf_df.pca_cor = princomp(num_leaf_df, cor=T)

pcasum = summary(num_leaf_df.pca_cor, loadings=T, cutoff=0.30)
pcasum



cum = c(0.4060298, 0.7055398, 0.8557138, 0.90826871, 0.9395322, 0.96733781, 0.97955083, 0.987623892, 0.992861029, 0.996111212, 0.997869154, 0.999118308, 0.9999826454, 1.000000)
barplot(cum*100, width = 0.5, space = 0.1, main="Variability Explained", axes = T)


```



##MDS Analysis
```{r}
leaf.dist = dist(num_leaf_df)

num_leaf_df.shape_feat

num_leaf_df.mds = cmdscale(leaf.dist, eig = TRUE)
num_leaf_df.mds

```
##Hierarchical Clustering
```{r}

agg = aggregate(full_df, list(full_df$Class), mean)

agg.shape_feat = agg[c(1:7)]

num_leaf_df.shape_feat = num_leaf_df[c(1:7)]
num_leaf_df.text_feat = num_leaf_df[c(8:14)]

fd = dist(num_leaf_df.shape_feat)

## Single linkage hierarchical clustering
hcs <- hclust(fd, "single")
plot(hcs, main = "Single Linkage HC Dendrogram")


plot(hcs, main = "Single Linkage HC Dendrogram")
plot(rev(hcs$height))
ct <- cutree(hcs, 12)
table(ct)

##complete linkage hierarchical clustering
hcc <- hclust(fd, "complete")
plot(hcc, main = "Complete Linkage HC Dendrogram")
plot(rev(hcc$height))
ct1 <- cutree(hcc, 7)
table(ct1)

##Average linkage hierarchical clustering
hca <- hclust(fd, "average")
plot(hca, main = "Average Linkage HC Dendrogram")
plot(rev(hca$height))
ct2 <- cutree(hca, 10)
table(ct2)


```



##Hierarchical Clustering
```{r}
# ## Single linkage hierarchical clustering
# hcs <- hclust(leaf.dist, "single")
# plot(hcs, main = "Single Linkage HC Dendrogram")
# plot(rev(hcs$height))
# ct <- cutree(hcs, 12)
# table(ct)
# 
# ##complete linkage hierarchical clustering
# hcc <- hclust(leaf.dist, "complete")
# plot(hcc, main = "Complete Linkage HC Dendrogram")
# plot(rev(hcc$height))
# ct1 <- cutree(hcc, 7)
# table(ct1)
# 
# ##Average linkage hierarchical clustering
# hca <- hclust(leaf.dist, "average")
# plot(hca, main = "Average Linkage HC Dendrogram")
# plot(rev(hca$height))
# ct2 <- cutree(hca, 10)
# table(ct2)


```
```{r}
######K-means
agg.shape_feat = agg[c(1:7)]
agg.text_feat = agg[c(8:14)]

num_leaf_df.shape_feat = num_leaf_df[c(1:7)]
num_leaf_df.text_feat = num_leaf_df[c(8:14)]

leaf.km <- kmeans(agg.text_feat, centers = 9, nstart = 10)
leaf.km$tot.withinss
table(leaf.km$cluster)

plot.wgss(num_leaf_df, 20)

leaf.km1 <- kmeans(num_leaf_df, centers = 5, nstart = 20)
leaf.km1$tot.withinss
table(leaf.km1$cluster)

plot(num_leaf_df.pca_cor$scores[,1:2], col = leaf.km1$cluster)


```



```{r}
######K-means

leaf.km <- kmeans(num_leaf_df, centers = 9, nstart = 10)
leaf.km$tot.withinss
table(leaf.km$cluster)

plot.wgss <- function(mydata, maxc){
  wss <- numeric(maxc)
  for (i in 1:maxc){
    wss[i] <- kmeans(mydata, iter.max = 100,
                     centers = i, nstart = 10)$tot.withinss
  }
  plot(1:maxc, wss, type = "b",
       xlab = "Number of Clusters",
       ylab = "Within Groups Sum of Squares",
       main = "Scree Plot")
}
plot.wgss(num_leaf_df, 20)

leaf.km1 <- kmeans(num_leaf_df, centers = 14, nstart = 10)
leaf.km1$tot.withinss
table(leaf.km1$cluster)

plot(num_leaf_df.pca_cor$scores[,1:2], col = leaf.km1$cluster)


```

```{r}
###Model-based clustering

library(mclust)
mbc.i <- Mclust(num_leaf_df)
table(mbc.i$classification)#we have seven clusters
plot(mbc.i, what = "BIC")
mbc.i$modelName


```


##CFA
```{r}


r.fa.2varimax =  factanal(num_agg, factors = 3)
print(r.fa.2varimax$loadings, cut = 0.5)

```

